### 2018-08-08 sklearn_pyTorch_baxter.Rmd
added pytorch code

---
title: "Logistic Regression, MLP and Random Forest with Sklearn and Feed Forward Neural Net with PyTorch on Baxter Dataset"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
library(reticulate)
use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = TRUE)
py_config()

```

#### Let's make sure we are using the Python version we want
```{python,tidy = TRUE, comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import sys
print(sys.version)
```
#### Let's make sure we are in our project directory
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import os
print(os.getcwd())
```
#### Start by importing modules that will be necessary 
```{python,tidy = TRUE, echo=T, results='hide', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
###### IMPORT MODULES #######
# We need to use a backend for matplotlib
# https://stackoverflow.com/questions/21784641/installation-issue-with-matplotlib-python/21789908#21789908
import matplotlib as mpl
mpl.use('TkAgg')
import matplotlib.pyplot as plt
from sklearn.linear_model import LogisticRegression
from sklearn.model_selection import train_test_split
from sklearn.model_selection import KFold, cross_val_score, validation_curve
from sklearn import linear_model
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
from sklearn.model_selection import GridSearchCV
from sklearn.neural_network import MLPClassifier
from scipy import interp
from itertools import cycle
from sklearn import svm, datasets
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import StratifiedKFold
# dependencies for statistic analysis
from scipy import stats
#importing our parameter tuning dependencies
from sklearn.neighbors import KNeighborsClassifier
from sklearn.model_selection import (cross_val_score, GridSearchCV, StratifiedKFold, ShuffleSplit )
#importing our dependencies for Feature Selection
from sklearn.feature_selection import (SelectKBest, chi2, RFE, RFECV)
from sklearn.linear_model import LogisticRegression, RandomizedLogisticRegression
from sklearn.ensemble import RandomForestClassifier
from sklearn.metrics import f1_score
# Importing our sklearn dependencies for the modeling
from sklearn.ensemble import RandomForestClassifier
from sklearn.neighbors import NearestNeighbors
from sklearn.preprocessing import StandardScaler
from sklearn.model_selection import train_test_split
from sklearn import metrics
from sklearn.metrics import (accuracy_score, confusion_matrix, classification_report, roc_curve, auc)
from sklearn.neural_network import MLPClassifier
from itertools import cycle
from scipy import interp
import warnings
from sklearn.model_selection import StratifiedKFold
import torch
from torch.autograd import Variable
import torch.utils.data as data_utils
import torch.nn.init as init
import torch.nn as nn
import torch.nn.functional as F
from torch.utils.data import Dataset, DataLoader
from sklearn.model_selection import train_test_split
from sklearn.preprocessing import StandardScaler
warnings.filterwarnings('ignore')
```
#### Read and prepare data 
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
## Read in the data
shared = pd.read_table("data/baxter.0.03.subsample.shared")
shared.head()
meta = pd.read_table("data/metadata.tsv")

## Check and visualize the data
print(meta.head())
print(shared.head())

## Remove unnecessary columns from meta
meta = meta[['sample','dx']]

## Rename the column name "Group" to match the "sample" in meta
shared = shared.rename(index=str, columns={"Group":"sample"})

## Merge the 2 datasets on sample
data=pd.merge(meta,shared,on=['sample'])

## Remove adenoma samples
data= data[data.dx.str.contains("adenoma") == False]

## Drop all except OTU columns for x
x = data.drop(["sample", "dx", "numOtus", "label"], axis=1)
## Cancer =1 Normal =0
diagnosis = { "cancer":1, "normal":0}
##Generate y which only has diagnosis as 0 and 1
y = data["dx"].replace(diagnosis)
# y = np.eye(2, dtype='uint8')[y]
## Drop if NA elements
y.dropna()
x.dropna()
print(x.head())
print(y.head())
print(len(x))
print(len(y))
```
#### Split the data to generate training and testing set %80-20
Here we also want to shuffle the data and set a random state
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, random_state=82089, shuffle=True)
```
#### Now let's define a L2 regularized logistic regression model
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

## Define L2 regularized logistic classifier
logreg = linear_model.LogisticRegression(C=0.01)

```
#### Plot ROC curve for Logistic Regression training set
We split the training set to 10 to cross-validate
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

cv = StratifiedKFold(n_splits=10)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)

## Convert from pandas dataframe to numpy
X=x_train.values
Y=y_train.values

# Plot the ROC curve over 10 iterations for each split
#logistic_plot = plt.figure()
i = 0
for train, test in cv.split(X,Y):
    probas_ = logreg.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
    fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
    tprs.append(interp(mean_fpr, fpr, tpr))
    tprs[-1][0] = 0.0
    roc_auc = auc(fpr, tpr)
    aucs.append(roc_auc)
    plt.plot(fpr, tpr, lw=1, alpha=0.3, label='ROC fold %d (AUC = %0.2f)' % (i, roc_auc))
    i += 1

plt.plot([0, 1], [0, 1], linestyle='--', color='r', label='Luck', alpha=.8)
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='grey', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('Logistic Regression ROC\n')
plt.legend(loc="lower right", fontsize=6)
plt.show()
#logistic_plot.savefig('results/figures/Logit_Baxter.png', dpi=1000)
```

#### Predict using the Logistic Regression model on the test set

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
y_pred = logreg.predict(x_test)
print('Accuracy of logistic regression classifier on test set: {:.2f}'.format(logreg.score(x_test, y_test)*100), "%")
```

#### Now let's define a Multi-layer Perceptron Neural Network Model

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
clf = MLPClassifier(activation='logistic', alpha=0.001, batch_size='auto',
       beta_1=0.9, beta_2=0.999, early_stopping=True, epsilon=1e-08,
       hidden_layer_sizes=(100,), learning_rate='adaptive',
       learning_rate_init=0.001, max_iter=200, momentum=0.9,
       nesterovs_momentum=True, power_t=0.5, random_state=1, shuffle=True,
       solver='sgd', tol=0.0001, validation_fraction=0.1, verbose=False,
       warm_start=False)

```
#### Plot ROC curve for MLP on training set
We split the training set to 10 to cross-validate
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

cv = StratifiedKFold(n_splits=10)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
mlp_plot = plt.figure()

i = 0
for train, test in cv.split(X,Y):
    probas_ = clf.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
    fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
    tprs.append(interp(mean_fpr, fpr, tpr))
    tprs[-1][0] = 0.0
    roc_auc = auc(fpr, tpr)
    aucs.append(roc_auc)
    plt.plot(fpr, tpr, lw=1, alpha=0.3, label='ROC fold %d (AUC = %0.2f)' % (i, roc_auc))
    i += 1

plt.plot([0, 1], [0, 1], linestyle='--', color='r', label='Luck', alpha=.8)
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='grey', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('Neural Network ROC\n')
plt.legend(loc="lower right", fontsize=7)
plt.show()
#mlp_plot.savefig('results/figures/MLP_Baxter.png', dpi=1000)
```

#### Predict using the MLP model on the test set

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
y_pred = clf.predict(x_test)
## Print accuracy
print("Performance Accuracy on the Testing data:", round(clf.score(x_test, y_test) *100))
print("Number of correct classifiers:", round(accuracy_score(y_test, y_pred, normalize=False)))
```
#### Now let's define a Random Forest model

```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

rfc = RandomForestClassifier(bootstrap=True, class_weight=None, criterion='gini',
            max_depth=None, max_features='auto', max_leaf_nodes=None,
            min_impurity_split=None, min_samples_leaf=1,
            min_samples_split=2, min_weight_fraction_leaf=0.0,
            n_estimators=150, n_jobs=1, oob_score=True, random_state=None,
            verbose=0, warm_start=False)
```
#### Plot ROC curve for Random Forest on training set
We split the training set to 10 to cross-validate
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

RF_plot = plt.figure()
cv = StratifiedKFold(n_splits=10)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)

i = 0
for train, test in cv.split(X,Y):
    probas_ = rfc.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
    fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
    tprs.append(interp(mean_fpr, fpr, tpr))
    tprs[-1][0] = 0.0
    roc_auc = auc(fpr, tpr)
    aucs.append(roc_auc)
    plt.plot(fpr, tpr, lw=1, alpha=0.3, label='ROC fold %d (AUC = %0.2f)' % (i, roc_auc))
    i += 1

plt.plot([0, 1], [0, 1], linestyle='--', color='r', label='Luck', alpha=.8)
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='grey', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('Random Forest ROC\n')
plt.legend(loc="lower right", fontsize=7)
plt.show()
#RF_plot.savefig('results/figures/Random_Forest_Baxter.png', dpi=1000)
```

#### Predict using the Random Forest model on the test set

```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
y_pred = rfc.predict(x_test)
print("Performance Accuracy on the Testing data:", round(rfc.score(x_test, y_test) *100))
print("Number of correct classifiers:", round(accuracy_score(y_test, y_pred, normalize=False)))
```

#### Now let's define a 1 hidden layer(n=100) Feed Forward Neural Net
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

class Net(nn.Module):
    def __init__(self):
        super(Net, self).__init__()
        self.fc1 = nn.Linear(6920, 100)
        self.fc2 = nn.Linear(100, 2)
    def forward(self, x):
        x = self.fc1(x)
        x = F.dropout(x, p=0.1)
        x = F.relu(x)
        x = self.fc2(x)
        x = F.sigmoid(x)
        return x

net = Net()

## Batch size allows for random sampling of the dataset during training
batch_size = 50
num_epochs = 50
learning_rate = 0.0001
batch_no = len(x_train) // batch_size

criterion = nn.CrossEntropyLoss()
optimizer = torch.optim.Adam(net.parameters(), lr=learning_rate)

from sklearn.utils import shuffle
from torch.autograd import Variable
from scipy import interp
from sklearn.metrics import (accuracy_score, confusion_matrix, classification_report, roc_curve, auc)

## ROC plot for training
pyTorch_plot = plt.figure()
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
for epoch in range(num_epochs):
    x_train, y_train = shuffle(x_train, y_train)
    # Mini batch learning
    for i in range(batch_no):
        start = i * batch_size
        end = start + batch_size
        x_var = Variable(torch.FloatTensor(x_train.values[start:end]))
        y_var = Variable(torch.LongTensor(y_train.values[start:end]))
        # Forward + Backward + Optimize
        ypred_var = net(x_var)
        loss =criterion(ypred_var, y_var)
        correct_num = 0
        ## The outputs of the model (ypred_var) are energies for the 10 classes. Higher the energy for a class, the more the network thinks that the image is of the particular class. So, let’s get the index of the highest energy:
        values, labels = torch.max(ypred_var, 1)
        correct_num = np.sum(labels.data.numpy() == y_var.numpy())
        fpr, tpr, thresholds = roc_curve(y_var.numpy(), labels.data.numpy())
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='ROC fold %d (AUC = %0.2f)' % (epoch, roc_auc))
        optimizer.zero_grad()
        loss.backward()
        optimizer.step()
        print('Epoch [%d], Loss:%.4f, Accuracy:%.4f' % (epoch, loss.data[0], correct_num/len(labels)))

plt.plot([0, 1], [0, 1], linestyle='--', color='r', label='Luck', alpha=.8)
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='grey', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('PyTorch Neural Network ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#pyTorch_plot.savefig('results/figures/pyTorch_Baxter.png', dpi=1000)

# Evaluate the model on test set
net.eval()
pred = net(torch.from_numpy(x_test.values).float())
pred = torch.max(pred,1)[1]
len(pred)
pred = pred.data.numpy()
print(accuracy_score(y_test, pred))
print(confusion_matrix(y_test, pred))
```