### 2018-08-13 baxter.Rmd
first commit

---
title: "Logistic Regression, Multi-layer Perceptron Neural Network, Random Forest, XGBoost and PyTorchon Baxter Dataset"
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
fontsize: 10pt
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(reticulate)
use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = TRUE)
py_config()

```

#### Let's make sure we are using the Python version we want
```{python,tidy = TRUE, comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import sys
print(sys.version)
```
#### Let's make sure we are in our project directory
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import os
print(os.getcwd())
```
#### Start by importing modules that will be necessary 
```{python,tidy = TRUE, include=FALSE, echo=F, results='hide', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
###### IMPORT MODULES #######
# We need to use a backend for matplotlib
# https://stackoverflow.com/questions/21784641/installation-issue-with-matplotlib-python/21789908#21789908
import matplotlib as mpl
mpl.use('TkAgg')
import matplotlib.pyplot as plt
from sklearn.linear_model import LogisticRegression
from sklearn.model_selection import train_test_split
from sklearn.model_selection import KFold, cross_val_score, validation_curve
from sklearn import linear_model
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
from sklearn.model_selection import GridSearchCV
from sklearn.neural_network import MLPClassifier
from scipy import interp
from itertools import cycle
from sklearn import svm, datasets
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import StratifiedKFold
# dependencies for statistic analysis
from scipy import stats
#importing our parameter tuning dependencies
from sklearn.neighbors import KNeighborsClassifier
from sklearn.model_selection import (cross_val_score, GridSearchCV, StratifiedKFold, ShuffleSplit )
#importing our dependencies for Feature Selection
from sklearn.feature_selection import (SelectKBest, chi2, RFE, RFECV)
from sklearn.linear_model import LogisticRegression, RandomizedLogisticRegression
from sklearn.ensemble import RandomForestClassifier
from sklearn.metrics import f1_score
# Importing our sklearn dependencies for the modeling
from sklearn.ensemble import RandomForestClassifier
from sklearn.neighbors import NearestNeighbors
from sklearn.preprocessing import StandardScaler
from sklearn.model_selection import train_test_split
from sklearn import metrics
from sklearn.metrics import (accuracy_score, confusion_matrix, classification_report, roc_curve, auc)
from sklearn.neural_network import MLPClassifier
from itertools import cycle
from scipy import interp
import warnings
from sklearn.model_selection import StratifiedKFold
import torch
from torch.autograd import Variable
import torch.utils.data as data_utils
import torch.nn.init as init
import torch.nn as nn
import torch.nn.functional as F
from torch.utils.data import Dataset, DataLoader
from sklearn.model_selection import train_test_split
from sklearn.preprocessing import StandardScaler
import xgboost
from xgboost import XGBClassifier
from sklearn.model_selection import train_test_split
from sklearn.grid_search import GridSearchCV, RandomizedSearchCV
from sklearn.datasets import make_classification
from sklearn.cross_validation import StratifiedKFold
import xgboost as xgb
import xgboost
from xgboost import XGBClassifier
from sklearn.model_selection import train_test_split
from sklearn.grid_search import GridSearchCV, RandomizedSearchCV
from sklearn.datasets import make_classification
from sklearn.cross_validation import StratifiedKFold
from sklearn.externals import joblib
from sklearn.metrics import precision_score
from sklearn import preprocessing
from sklearn import cross_validation
from sklearn.model_selection import KFold
from sklearn.linear_model import LogisticRegression
from sklearn.model_selection import train_test_split
from sklearn.model_selection import KFold, cross_val_score, validation_curve
from sklearn import linear_model
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
import matplotlib.pyplot as plt
from sklearn.model_selection import GridSearchCV
from sklearn.neural_network import MLPClassifier
import numpy as np
from scipy import interp
from itertools import cycle
from sklearn import svm, datasets
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import StratifiedKFold
from pylab import rcParams
# dependencies for statistic analysis
from scipy import stats
#importing our parameter tuning dependencies
from sklearn.neighbors import KNeighborsClassifier
from sklearn.model_selection import (cross_val_score, GridSearchCV, StratifiedKFold, ShuffleSplit )
#importing our dependencies for Feature Selection
from sklearn.feature_selection import (SelectKBest, chi2, RFE, RFECV)
from sklearn.linear_model import LogisticRegression, RandomizedLogisticRegression
from sklearn.ensemble import ExtraTreesClassifier
from sklearn.cross_validation import ShuffleSplit
from sklearn.ensemble import RandomForestClassifier
from sklearn.metrics import f1_score
from collections import defaultdict
# Importing our sklearn dependencies for the modeling
from sklearn.ensemble import RandomForestClassifier
from sklearn.neighbors import NearestNeighbors
from sklearn.preprocessing import StandardScaler
from sklearn.cross_validation import KFold
from sklearn.model_selection import train_test_split
from sklearn import metrics
from sklearn.metrics import (accuracy_score, confusion_matrix, classification_report, roc_curve, auc)
from sklearn.neural_network import MLPClassifier
from itertools import cycle
from scipy import interp
import warnings
warnings.filterwarnings('ignore')
```
#### Read and prepare data 
We will use OTUs for each sample as our features and diagnosis column in the metadata as our classifier.
Cancer is 1 and Normal is 0.

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
## Read in the data
shared = pd.read_table("data/baxter.0.03.subsample.shared")
shared.head()
meta = pd.read_table("data/metadata.tsv")

## Check and visualize the data
print(meta.head())
print(shared.head())

## Remove unnecessary columns from meta
meta = meta[['sample','dx']]

## Rename the column name "Group" to match the "sample" in meta
shared = shared.rename(index=str, columns={"Group":"sample"})

## Merge the 2 datasets on sample
data=pd.merge(meta,shared,on=['sample'])

## Remove adenoma samples
data= data[data.dx.str.contains("adenoma") == False]

## Drop all except OTU columns for x
x = data.drop(["sample", "dx", "numOtus", "label"], axis=1)
## Cancer =1 Normal =0
diagnosis = { "cancer":1, "normal":0}
##Generate y which only has diagnosis as 0 and 1
y = data["dx"].replace(diagnosis)
# y = np.eye(2, dtype='uint8')[y]
## Drop if NA elements
y.dropna()
x.dropna()
print(x.head())
print(y.head())
print(len(x))
print(len(y))
```
#### Now let's define a L2 regularized logistic regression model
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

## Define L2 regularized logistic classifier
logreg = linear_model.LogisticRegression(C=0.01)

```
#### Plot ROC curve for Logistic Regression model on training set and predict test set
We split training and test set 80-20%. We further split the training set to 5 to cross-validate in a loop.
We predict test set in the same loop. 
We generate mean AUC for cross-validation with training set upon 50 epochs and mean AUC for test set upon 50 epochs. 
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

cv = StratifiedKFold(n_splits=5)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

# Plot the ROC curve over 10 iterations for each split
Logit_plot = plt.figure()
epochs = 50
for epoch in range(epochs):
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, shuffle=True)
    X=x_train.values
    Y=y_train.values
    X_test= x_test.values
    Y_test= y_test.values
    i = 0
    for train, test in cv.split(X,Y):
        probas_ = logreg.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
        fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='Epoch %d, ROC fold %d (AUC = %0.2f)' % (epoch, i, roc_auc))
        #i += 1
    probas_ = logreg.predict_proba(X_test)
    # Compute ROC curve and area the curve
    fpr_test, tpr_test, thresholds_test = roc_curve(Y_test, probas_[:, 1])
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Mean test ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('L2 Regularized Logistic Regression ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#Logit_plot.savefig('results/figures/Logit_Baxter.png', dpi=1000)
```

#### Now let's define a Multi-layer Perceptron Neural Network Model
Here the activation function is a logistic regression. There is 1 hidden layer with 100 nodes. 
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
clf = MLPClassifier(activation='logistic', alpha=0.001, batch_size='auto',
       beta_1=0.9, beta_2=0.999, early_stopping=True, epsilon=1e-08,
       hidden_layer_sizes=(100,), learning_rate='adaptive',
       learning_rate_init=0.001, max_iter=200, momentum=0.9,
       nesterovs_momentum=True, power_t=0.5, random_state=1, shuffle=True,
       solver='sgd', tol=0.0001, validation_fraction=0.1, verbose=False,
       warm_start=False)

```
#### Plot ROC curve for MLP on training set
We split training and test set 80-20%. We further split the training set to 5 to cross-validate in a loop.
We predict test set in the same loop. 
We generate mean AUC for cross-validation with training set upon 50 epochs and mean AUC for test set upon 50 epochs. 
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

cv = StratifiedKFold(n_splits=5)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

MLP_plot = plt.figure()
epochs = 50
for epoch in range(epochs):
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, shuffle=True)
    X=x_train.values
    Y=y_train.values
    X_test= x_test.values
    Y_test= y_test.values
    i = 0
    for train, test in cv.split(X,Y):
        probas_ = clf.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
        fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='Epoch %d, ROC fold %d (AUC = %0.2f)' % (epoch, i, roc_auc))
        #i += 1
    probas_ = clf.predict_proba(X_test)
    # Compute ROC curve and area the curve
    fpr_test, tpr_test, thresholds_test = roc_curve(Y_test, probas_[:, 1])
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Mean test ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('Multi-layer Perceptron Neural Network Model ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#MLP_plot.savefig('results/figures/MLP_Baxter.png', dpi=1000)
```

#### Now let's define a Random Forest model
Previous cross validation step allowed us to choose n_estimators(decision trees). Here we have 150 decision trees.
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

rfc = RandomForestClassifier(bootstrap=True, class_weight=None, criterion='gini',
            max_depth=None, max_features='auto', max_leaf_nodes=None,
            min_impurity_split=None, min_samples_leaf=1,
            min_samples_split=2, min_weight_fraction_leaf=0.0,
            n_estimators=150, n_jobs=1, oob_score=True, random_state=None,
            verbose=0, warm_start=False)
```
#### Plot ROC curve for Random Forest on training set and predict test set
We split training and test set 80-20%. We further split the training set to 5 to cross-validate in a loop.
We predict test set in the same loop. 
We generate mean AUC for cross-validation with training set upon 50 epochs and mean AUC for test set upon 50 epochs. 
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

RF_plot = plt.figure()
epochs = 50
for epoch in range(epochs):
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, shuffle=True)
    X=x_train.values
    Y=y_train.values
    X_test= x_test.values
    Y_test= y_test.values
    i = 0
    for train, test in cv.split(X,Y):
        probas_ = rfc.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
        fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='Epoch %d, ROC fold %d (AUC = %0.2f)' % (epoch, i, roc_auc))
        #i += 1
    probas_ = rfc.predict_proba(X_test)
    # Compute ROC curve and area the curve
    fpr_test, tpr_test, thresholds_test = roc_curve(Y_test, probas_[:, 1])
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Mean test ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('Random Forest ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#RF_plot.savefig('results/figures/RF_Baxter.png', dpi=1000)
```

#### Now let's define a XGBoost model
Previous cross validation step allowed us to choose n_estimators(decision trees), max_depth and learning rate.
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

clf = xgb.XGBClassifier(base_score=0.5, booster='gbtree', colsample_bylevel=1,
       colsample_bytree=1, gamma=0, learning_rate=0.01, max_delta_step=0,
       max_depth=1, min_child_weight=1, missing=None, n_estimators=250,
       n_jobs=1, nthread=None, objective='binary:logistic', random_state=0,
       reg_alpha=0, reg_lambda=0.1, scale_pos_weight=1, seed=None,
       silent=True, subsample=1)
```
#### Plot ROC curve for XGBoost on training set and predict test set
We split training and test set 80-20%. We further split the training set to 5 to cross-validate in a loop.
We predict test set in the same loop. 
We generate mean AUC for cross-validation with training set upon 50 epochs and mean AUC for test set upon 50 epochs. 
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

cv = StratifiedKFold(n_splits=5)
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

XGboost_plot = plt.figure()
epochs = 50
for epoch in range(epochs):
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, shuffle=True)
    X=x_train.values
    Y=y_train.values
    X_test= x_test.values
    Y_test= y_test.values
    i = 0
    for train, test in cv.split(X,Y):
        probas_ = clf.fit(X[train], Y[train]).predict_proba(X[test])
    # Compute ROC curve and area the curve
        fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='Epoch %d, ROC fold %d (AUC = %0.2f)' % (epoch, i, roc_auc))
        #i += 1
    probas_ = clf.predict_proba(X_test)
    # Compute ROC curve and area the curve
    fpr_test, tpr_test, thresholds_test = roc_curve(Y_test, probas_[:, 1])
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Mean test ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('XGBoost ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#XGboost_plot.savefig('results/figures/XGBoost_Baxter.png', dpi=1000)
```

#### Now let's define a 1 hidden layer(n=100) Feed Forward Neural Net
PyTorch allows for forward computation. One epoch will mean getting the output values(the way our model defines them),  one backward pass where weights in the model are updated and then optimize the hyperparameters for the next epoch.
```{python,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}

class Net(nn.Module):
    def __init__(self):
        super(Net, self).__init__()
        self.fc1 = nn.Linear(6920, 100)
        self.fc2 = nn.Linear(100, 2)
    def forward(self, x):
        x = self.fc1(x)
        x = F.dropout(x, p=0.1)
        x = F.relu(x)
        x = self.fc2(x)
        x = F.sigmoid(x)
        return x

net = Net()

## Batch size allows for random sampling of the dataset during training
batch_size = 50
num_epochs = 40
learning_rate = 0.0001
batch_no = 233 // batch_size

# Define Loss function and optimizer function
criterion = nn.CrossEntropyLoss()
optimizer = torch.optim.Adam(net.parameters(), lr=learning_rate)

## Plot ROC for training (with minibatches) and evaluate test sets
pyTorch_plot = plt.figure()
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

for epoch in range(num_epochs):
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2, shuffle=True)
    # Mini batch learning
    for i in range(batch_no):
        start = i * batch_size
        end = start + batch_size
        x_var = Variable(torch.FloatTensor(x_train.values[start:end]))
        y_var = Variable(torch.LongTensor(y_train.values[start:end]))
        # Forward + Backward + Optimize
        ypred_var = net(x_var)
        loss =criterion(ypred_var, y_var)
        correct_num = 0
        ## The outputs of the model (ypred_var) are energies for the 10 classes. Higher the energy for a class, the more the network thinks that the image is of the particular class. So, let’s get the index of the highest energy:
        values, labels = torch.max(ypred_var, 1)
        correct_num = np.sum(labels.data.numpy() == y_var.numpy())
        fpr, tpr, thresholds = roc_curve(y_var.numpy(), labels.data.numpy())
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        optimizer.zero_grad()
        loss.backward()
        optimizer.step()
        print('Epoch [%d], Loss:%.4f, Accuracy:%.4f' % (epoch, loss.data[0], correct_num/len(labels)))
        #plt.plot(fpr, tpr, lw=1, alpha=0.3, label='ROC fold %d (AUC = %0.2f)' % (epoch, roc_auc))
   #Predict test set and plot mean AUC
    x_var_test = Variable(torch.FloatTensor(x_test.values))
    y_var_test = Variable(torch.LongTensor(y_test.values))
    net.eval()
    ypred_var_test = net(x_var_test)
    correct_num_test = 0
    values_test, labels_test = torch.max(ypred_var_test, 1)
    correct_num_test = np.sum(labels_test.data.numpy() == y_var_test.numpy())
    fpr_test, tpr_test, thresholds_test = roc_curve(y_var_test.numpy(), labels_test.data.numpy())
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Mean test ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('PyTorch 1 hidden layer with 100 nodes Neural Network ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
```