### 2018-08-27 LogisticRegression_BaxterData.Rmd
submit to Jenna

---
title: "Logistic Regression on Baxter colorectal cancer dataset"
output: html_document

---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(reticulate)
use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = TRUE)
py_config()
```

#### Let's make sure we are using the Python version we want
```{python,tidy = TRUE, comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import sys
print(sys.version)
```

#### Let's make sure we are in our project directory
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import os
print(os.getcwd())
```

#### Start by importing modules that will be necessary 
```{python,tidy = TRUE, cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
############## IMPORT MODULES ######################
from sklearn.linear_model import LogisticRegression
from sklearn import linear_model
from sklearn.model_selection import train_test_split
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
import matplotlib.pyplot as plt
from scipy import interp
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import RepeatedStratifiedKFold
from sklearn.model_selection import GridSearchCV
```

#### Pre-process the data

We have 6920 OTUs that are the features we put in.
We have 292 samples with labels:
- 120 Cancer colon
- 172 Normal colon

I would like to do a classification model to predict discrete values (diagnosis 1 or 0)

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
############## PRE-PROCESS DATA ######################

## Read in the data

## shared has our input features(OTUs)
shared = pd.read_table("data/baxter.0.03.subsample.shared")
## meta has the labels(diagnosis)
meta = pd.read_table("data/metadata.tsv")

## Check and visualize the data
meta.head()
shared.head()

## Remove unnecessary columns from meta and only keep label for classification(diagnosis) and the sample name
meta = meta[['sample','dx']]

## Rename the column name "Group" to match the "sample" in meta
shared = shared.rename(index=str, columns={"Group":"sample"})

## Merge the 2 datasets on sample
data=pd.merge(meta,shared,on=['sample'])

## Remove adenoma samples. We will do a 2 classification model, just for cancer and normal colon samples.
data= data[data.dx.str.contains("adenoma") == False]

## Drop all except OTU columns for x
x = data.drop(["sample", "dx", "numOtus", "label"], axis=1)

## Cancer =1 Normal =0
diagnosis = { "cancer":1, "normal":0}

## Generate y which only has diagnosis as 0 and 1
y = data["dx"].replace(diagnosis)
# y = np.eye(2, dtype='uint8')[y]

## Drop if NA elements
y.dropna()
x.dropna()

## Split dataset to 80% training 20% test sets.

x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2,shuffle=True)

```

#### Now let's define a L2 regularized logistic regression model
- We tune our hyper-parameter(regualarization strength) here with 80-20 splits and 1000 repeats. 
- Once we have our "C" value tuned, we define our best model.
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
######################### Logistic Regression ##############################

## Define L2 regularized logistic classifier
logreg = linear_model.LogisticRegression()

## Define the n-folds for hyper-parameter optimization on training set.
cv = RepeatedStratifiedKFold(n_splits=5, n_repeats=1000, random_state=200889)
## We will try these regularization strength coefficients to optimize our model
C = {"C": [0.0000001, 0.000001, 0.00001, 0.0001]}
## Define the best model:
clf = GridSearchCV(logreg, C, cv=cv, verbose=0, scoring='roc_auc')
best_model = clf.fit(x_train, y_train)
print('Best C:', best_model.best_estimator_.get_params()['C'])
print('Best model:', best_model.best_estimator_)
## The best model we pick here will be used for predicting test set.
best_model = best_model.best_estimator_

```

#### Plot ROC curve for cross validation of training set and then test set 
1. Plot mean ROC curve for cross-validation with n_splits=5 (80-20 stratified) and n_repeats=1000 to evaluate the variation of prediction in our training set.
2. Plot ROC curve for never before seen test set.

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
## Now let's plot some ROC curves
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)

tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

Logit_plot = plt.figure()

## Converting to numpy array from pandas
X=x_train.values
Y=y_train.values
X_test= x_test.values
Y_test= y_test.values

## Plot mean ROC curve for cross-validation with n_splits=5 and n_repeats=1000 to evaluate the variation of prediction in our training set.
for train, test in cv.split(X,Y):
    probas_ = best_model.fit(X[train], Y[train]).predict_proba(X[test])
    fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
    tprs.append(interp(mean_fpr, fpr, tpr))
    tprs[-1][0] = 0.0
    roc_auc = auc(fpr, tpr)
    aucs.append(roc_auc)

## Predict and plot ROC curve for never before seen test set.
probas_ = best_model.predict_proba(x_test)
# Compute ROC curve and area the curve
fpr_test, tpr_test, thresholds_test = roc_curve(y_test, probas_[:, 1])
tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
tprs_test[-1][0] = 0.0
roc_auc_test = auc(fpr_test, tpr_test)
aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Never-before-seen test set ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean cross-val ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('L2 Logistic Regression ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#Logit_plot.savefig('results/figures/Logit_Baxter.png', dpi=1000)

```