### 2018-09-19 results/ML_outline.Rmd
shared with Jenna

---
title: "Machine Learning Manuscript Outline"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

##### General Context of the work
As gut microbiome field continues to grow, there will be an ever-increasing demand for reproducible machine learning  methods to analyze microbiome sequence read count data and to determine association with a continuous or categorical phenotype of interest.

##### Narrower research area and statement of its importance
* Colorectal cancer is one of the leading cause of death among cancers in the United States. Early diagnosis increases the chance of survival. However the current diagnostic methods are expensive and invasive. As a less invasive tool, numerous studies use relative abundances of the gut bacteria populations to predict disease progression. 
* Most microbial communities are pretty patchy and the likelihood of a single feature that explains the differences in health is pretty small. It is likely that many biomarkers are needed to account for the patchiness as well as the context dependency of the features.

##### Identification of a gap or other need for research
* ML use in microbiome literature is a bit like the wild west with lack of clarity over methods, testing, validation, etc. There is a need for guidance on how to properly implement these different methods.
* We need to emphasize good machine learning practices and pipelines and discuss the reproducibility, robustness and actionability of models.

##### Summary of appraoch to answer the research question
* We establish a non-leaky pipeline. 
* We perform L1 and L2-regularized logistic regression, Linear SVM, Non-Linear SVM, Decision tree, Random forest, XGBoost and Feed Forward Neural Net classification models.
* We evaluate the classification success of different machine learning methods. We also want to discuss the reproducibility, robustness, actionability, interpretibility and susceptibility to overfitting of each method.

##### Announcement of principal findings
* Generalisation Perfomance of each model. 
* Is there a maximum threshold of prediction with all these methods?
* Does an increase in model complexity improve predictibility?
* Synthesis statement regarding modeling 16S microbiome data

## Methods
##### Brief explanation of study design/patient sampling and 16S rRNA gene sequencing/curation
* Refer to Baxter Dataset:
We sequenced the 16S rRNA genes from the stool samples of 292 patients, 120 with colorectal cancer
and 172 with healthy colons. We used the relative abundances of the bacterial populations within each sample to develop classification models that detects colonic lesions using the relative abundance of gut microbiota.

##### Analysis of data
* What is the data (temporal or not)? Assumptions we make when we use a dataset. What will the future data look like?
* Pre-proccessing of the data
* Machine Learning pipeline backbone. How do I split, train, validate and test?
* Cross-validation and hyper-parameter tuning methods for each modeling method
* Programming languages and packages/modules utilized
* Statistical methods for comparison of model performance
* There is likely to be a tradeoff between interpretability and performance (perhaps this should go in a paragraph about ML methods that would also indicate which methods are linear/non-linear.)

## Results
##### Results of modeling in text, tables and figures
##### Comparisons among modeling approaches

## Discussion
#### Interpretation of modeling results in terms of reproducibility, robustness, actionability, interpretibility and susceptibility
#### Consideration of possible weaknesses for each model 
* The interactions between the biomarkers may be nonlinear. Obviously, the linear models will not incorporate this because they are linear. Tools like linear models (e.g. metastats, lefse, wilcoxon, etc) are likely worthless.

#### Consideration of possible weaknesses for our approach and chosen dataset
#### Relationship of results to previous literature and broader implications of this work
#### Prospects of future progress