### 2018-11-09 results/tables/Table2.Rmd

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
---
```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

######################################################################
#-------------- Define knitr options to allow kable ----------------#
######################################################################
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
library(kableExtra)
```

```{r table_2, echo=FALSE, results='asis'}
######################################################################
# We load a param_grid.scv file to make our Table 2 and 3. This file has the information of the ranges of the optimized hyper-parameters for each model.

# This file is generated from model_selection.py that is used to define model parameters for main.py. Therefore any changes to those files would allow us to represent them in the tables automatically.

# We will generate a Table for non-tree based models that shows C(penalty) and gamma hyperparameters.
######################################################################

################ Non-tree based model parameters ##############
linear_param_range <- read.table('../../data/process/param_grid.csv', header=T, sep=',', na.strings="", stringsAsFactors = FALSE) %>% 
  replace(is.na(.), "-")  %>% 
  ## only select Logistic Regression and SVM model columns
  select(1:17) %>% 
  ## only select C (penalty) and gamma rows as parameters
   slice(1:2)

#------------------------Table 2 generation------------------------#
Table2 <- kable(linear_param_range, 
                "latex", 
                caption = "Table 2. The range of optimized hyper-parameters for logistic regression and support vector machines.", 
                booktabs = T, 
                col.names = NULL, 
                linesep = "") %>%
  # latex options make table to be condensed
  # table itseld if left aligned
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), 
                position = "left") %>% 
  # the column names are removed when the file is read so I put headers above the grouped/replicated modeling method for
  add_header_above(c("Parameter", "L2 Logistic" = 4, "L1 SVM Linear" = 4, "L2 SVM Linear " = 3, "SVM RBF"=5), bold=TRUE) %>% 
  # Put a left border to seperate above header lines
  column_spec(2, border_left = T) %>%
  column_spec(6, border_left = T) %>% 
  column_spec(10, border_left = T) %>% 
  column_spec(13, border_left = T)

Table2
```