### 2018-11-09 results/tables/Table3.Rmd

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
---
```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

######################################################################
#-------------- Define knitr options to allow kable ----------------#
######################################################################
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
library(kableExtra)
```

```{r table_3, echo=FALSE, results='asis'}
######################################################################
# We load a param_grid.scv file to make our Table 2 and 3. This file has the information of the ranges of the optimized hyper-parameters for each model.

# This file is generated from model_selection.py that is used to define model parameters for main.py. Therefore any changes to those files would allow us to represent them in the tables automatically.

# We will generate a Table for tree based models that show all the hyperparameters used for it. 
######################################################################

################ Tree based model parameters ##############
tree_param_range <- read.table('../../data/process/param_grid.csv', header=T, sep=',', na.strings="", stringsAsFactors = FALSE) %>% 
  replace(is.na(.), "-")  %>% 
  ## only select Logistic Regression and SVM model columns
  select(1, 18:29) %>% 
  ## select all rows except C (penalty) and gamma for parameters
  slice(3:9)

#------------------------Table 3 generation------------------------#
Table3 <- kable(tree_param_range, 
                "latex", 
                caption = "Table 3. The range of optimized hyper-parameters for tree based classification algorithms.", 
                booktabs = T, 
                col.names = NULL, 
                linesep = "") %>%
kable_styling(latex_options = c("striped", "scale_down", "hold_position"), 
              position = "left", font_size=6) %>% 
  add_header_above(c("Parameter", "Random Forest"=5, "Decision Tree"= 4, "XGBoost"=3 ), bold=TRUE) %>% 
  column_spec(2, border_left = T) %>% 
  column_spec(7, border_left = T)%>% 
  column_spec(11, border_left = T)

Table3
```