### 2018-11-14 results/tables/Table1.Rmd
fix commented code

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}

####################################################################################################
#------------------------------- Read in necessary libraries --------------------------------------#
####################################################################################################
deps = c("knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}

#####################################################################################################
#------------------------------- Define knitr options to allow kable ------------------------------#
####################################################################################################
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
library(kableExtra)
```

```{r table_1_models, echo=FALSE, results='asis'}

#####################################################################################################

# We load a model parameters file to make our Table 1. Model parameters file has the information of optimized hyper-parameters, pre-processing and cross-validation methods and software implementation for all the classification algorithms.

# This file is hand-made. Therefore needs to be changed manually if new information wants to be added or information needs to be changed.
##################################################################################################

model_param <- read.table('../../data/process/model_parameters.txt', header=T, sep='\t') %>%
  select(Method, Parameter, Cross.Validation, Epochs, Scaler, Sklearn.Function) %>%
  rename("Inner CV"="Cross.Validation") %>%
  rename("Outer CV"="Epochs") %>%
  rename("Function"="Sklearn.Function") %>%
  rename("Scaling"="Scaler") %>%
  rename("Hyper-Parameter"="Parameter")

#--------------------------------------- Table 1 generation ---------------------------------------#

Table1 <- kable(model_param,
                "latex",
                caption = "Table 1. Optimized hyper-parameters, pre-processing and inner and outer cross-validation (CV) methods and software implementation of the classification algorithms.",
                booktabs = T,
                linesep = "") %>%
kable_styling(latex_options = c("striped", "scale_down", "hold_position"),
              position = "left") %>%
 column_spec(1, width="10em") %>%
  column_spec(2, width="8em") %>%
  column_spec(3, width="9em") %>%
  column_spec(5, width="5em") %>%
  row_spec(1:7, align="l") %>%
  row_spec(0,bold=TRUE)

Table1
```