### 2018-11-28 submission/Microbe_Abstract2019.Rmd
microbe conference submission

---
bibliography: references.bib
output: word_document
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
geometry: margin=1.0in
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
######################################################################
#-------------- Define the chunk options ----------------#
######################################################################
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
library(reticulate)
use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = TRUE)
```

```{r LoadData, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
######################################################################
#----------------- Load OTU table and MetaData -----------------#
######################################################################

# Read in metadata and select only sample Id and diagnosis columns
meta <- read.delim('../data/metadata.tsv', header=T, sep='\t') %>%
  select(sample, dx, Dx_Bin, fit_result)

# Read in OTU table and remove label and numOtus columns
shared <- read.delim('../data/baxter.0.03.subsample.shared', header=T, sep='\t') %>%
   select(-label, -numOtus)

######################################################################
#---Load .tsv data with CV and test AUC data generated in Python ----#
######################################################################

# Read in AUCs table generated from L2 logistic regression model for all samples:
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax scaler
logit <- read.delim('../data/process/L2_Logistic_Regression.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-Logistic Regression")

logit_summary <- logit %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))

cv_meanAUC <- logit_summary[1,2]
cv_sdAUC <- logit_summary[1,3]
test_meanAUC <- logit_summary[2,2]
test_sdAUC <- logit_summary[2,3]

# Read in AUCs table generated from l1 SVM linear kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
l1svm <- read.delim('../data/process/L1_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L1-SVM Linear")

l1svm_summary <- l1svm %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))

# Read in AUCs table generated from l2 SVM linear kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
l2svm <- read.delim('../data/process/L2_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-SVM Linear")

l2svm_summary <- l2svm %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))

# Read in AUCs table generated from  SVM RBF kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
svmRBF <- read.delim('../data/process/SVM_RBF.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="SVM RBF")

svmRBF_summary <- svmRBF %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))

# Read in AUCs table generated from xgboost for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
xgboost <- read.delim('../data/process/XGBoost.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="XGBoost")

xgboost_summary <- xgboost %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))

# Read in AUCs table generated from random forest for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
rf <- read.delim('../data/process/Random_Forest.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Random Forest ")

rf_summary <- rf %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))
# Read in AUCs table generated from decision tree for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
dt <- read.delim('../data/process/Decision_Tree.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Decision Tree")

dt_summary <- dt %>%  
  group_by(Performance) %>% 
  summarise(meanAUC=mean(AUC), std=sd(AUC))
```

```{python, cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
import platform
python_version = platform.python_version()
import sklearn
sklearn_module = sklearn.__version__
```
\linenumbers
__Evaluation of machine learning methods that predict colorectal cancer progression with microbiota-associated biomarkers__

Begüm D. Topçuoğlu, Jenna Wiens, Mack Ruffin, Patrick D. Schloss

As gut microbiome field continues to grow, there will be an ever-increasing demand for reproducible machine learning methods to determine association of the microbiome with a continuous or categorical phenotype of interest. Currently, the use of machine learning in microbiome research lack clarity and consistency over the training, validation and testing of models. There is a need for guidance on how to properly implement good machine learning practices to generate reproducible and robust models.

Recently, there has been growing interest in using machine learning to predict colorectal cancer progression with microbiota-associated biomarkers. Colorectal cancer is one of the leading cause of death among cancers in the United States. Colonoscopy as a screening tool is effective, however it is invasive, expensive and have a low rate of patient adherence. Previous studies have shown that bacterial population abundances in the stool can predict screen relevant growth in the colon and can be used as a non-invasive screening tool. However, the prediction performance of these models vary greatly, with areas under the receiver operating characteristic curve (AUC) of 0.7-0.9 [@sze_leveraging_2018; @baxter_microbiota-based_2016; @baxter_dna_2016; @zackular_human_2014]. The variation in prediction performance is based in part on differences in the study populations, and in part on the differences in training, validation and testing pipelines.

In this study, hemoglobin levels and bacterial population abundances in the stool were used to predict colorectal disease status of `r paste(nrow(meta))` patients. The colorectal disease status was defined as showing screen-relevant colonic growth or not. Training, validation and testing pipelines were established for L2-regularized Logistic Regression, L1 and L2 Linear Support Vector Machines (SVM), Radial Basis Function (RBF) SVM, Decision Tree, Random Forest and XGBoost binary classifiers. L2-regularized Logistic Regression, L1 Linear SVM, L2 Linear SVM, and RBF SVM had mean AUCs of `r paste(round(test_meanAUC,2))` ± `r paste(round(test_sdAUC,2))`,  `r paste(round(l1svm_summary[2,2],2))` ± `r paste(round(l1svm_summary[2,3],2))`,  `r paste(round(l2svm_summary[2,2],2))` ± `r paste(round(l2svm_summary[2,3],2))`, and `r paste(round(svmRBF_summary[2,2],2))` ± `r paste(round(svmRBF_summary[2,3],2))`, respectively. Decision Tree, Random Forest and XGBoost had mean AUCs of `r paste(round(dt_summary[2,2],2))` ± `r paste(round(dt_summary[2,3],2))`, `r paste(round(rf_summary[2,2],2))` ± `r paste(round(rf_summary[2,3],2))`, `r paste(round(xgboost_summary[2,2],2))` ± `r paste(round(xgboost_summary[2,3],2))`, respectively. Tree-based models were less susceptible to overfitting and in general had higher sensitivity and specificity for colonic screen-relevant growth. Aside from evaluating generalization and classification performance of each classifier, this study established standards for training, validation and testing of the microbiome-associated machine learning models.

\newpage

__References__

<div id="refs"></div>