### 2019-06-26 submission/TableS1.Rmd
update

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
fontsize: 11pt
---

\begin{center}
\captionof{table}{Model complexity: $n$ is the number of training samples, $f$ the number of features, $n_{trees}$ is the number of trees.}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Machine learning model} & \textbf{Training complexity} & \textbf{Hyperparameters} \\ \hline
Logistic regression & $O(nf)$ & $C$ \\ \hline
SVM with linear kernel & $O(nf)$ & $C$ \\ \hline
SVM with radial basis kernel & $O(n^2f)$ & $C$, $\sigma$ \\ \hline
Decision tree & $O(n^2f)$ & tree depth \\ \hline
Random forest & $O(n^2fn_{trees})$ & number of features \\ \hline
XGBoost & $O(nfn_{trees})?$  \\ \hline
\end{tabular}
\end{center}