### 2019-07-18 submission/Table1.Rmd
add Table 1 and 2

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
fontsize: 11pt
---

\newpage
\captionof{table}{Characteristics of the machine learning models in our comparative study.}
\small
\begin{tabular}{|l|l|l|l|l|}
\hline

\rowcolor{lightgray}
\textbf{Model} & \textbf{Description} & \textbf{Linearity} & \textbf{Interpretability} & \textbf{Refs.} \\ \hline

\makecell[l]{Logistic \\regression} & \makecell[l]{A predictive regression analysis when the dependent \\variable is binary.} & Linear & Interpretable & $^{36}$\\ \hline

\makecell[l]{SVM with \\linear kernel} & \makecell[l]{A classifier that is defined by an optimal linear \\seperating hyperplane that discriminates between labels.} & Linear & Interpratable & $^{37}$\\ \hline

\makecell[l]{SVM with \\radial basis kernel} & \makecell[l]{A classifier that is defined by an optimal gaussian \\seperating hyperplane that discriminates between labels.} & Non-linear & Explainable$^*$ & $^{38}$\\ \hline
Decision tree & \makecell[l]{A classifier that sorts samples down from the 
root to the \\leaf node where an attribute is tested to discriminate \\between labels} & Non-linear & Interpretable & $^{39}$\\ \hline

Random forest & \makecell[l]{A classifier that is a decision tree ensemble that \\grow randomly with subsampled data.} & Non-linear & Explainable$^*$ & $^{40-41}$  \\ \hline

XGBoost & \makecell[l]{A classifier that is a decision tree ensemble that \\grow with additive training.} & Non-linear & Explainable$^*$ & $^{42-43}$ \\ \hline

\end{tabular}
\begin{tablenotes}\footnotesize
\item[1]{$^*$Explainable models are not inherently interpretable but can be explained with post-hoc analyses.}
\end{tablenotes}
\newpage